\textbf{\emph{De novo} clustering methods out-perform reference-based methods for assigning 16S rRNA gene sequences to operational taxonomic units}

### 2016-02-22 papers/peerj_2015/Schloss_Cluster_PeerJ_2015.tex
Move PeerJ paper files to papers/peerj_2015

\textbf{\emph{How do the various methods vary in the quality of their
OTU assignments?}} More important than the stability of OTUs is whether
sequences are assigned to the correct OTUs. A method can generate highly
stable OTUs, but the OTU assignments may be meaningless if they poorly
represent the specified cutoff and the actual distance between the
sequences. To assess the quality of OTU assignments by the various
methods, we made use of the pairwise distance between the unique
sequences to count the number of true positives and negatives and the
number of false positives and negatives for each method and sampling
depth. Those pairs of sequences that were similar to each other and
found in the same OTU were called true positives while those that were
similar and found in different OTUs were called false negatives.
Meanwhile, those pairs of sequences that were different from each other
and found in the same OTU were called false positives and those that
were dissimilar and found in different OTUs were called true negatives.
Counting the frequency of these different classes allowed us to judge
how each method balanced the ratio of true positives and negatives to
false positives and negatives using the MCC. We used the average MCC
value as a measure of a method's quality and its variation as a measure
of its consistency. We made three important observations. First, each of
the \emph{de novo} methods varied in how sensitive their MCC values were
to additional sequences (Figure 1C). The SL and CL methods were the most
sensitive; however, the quality of the OTU assignments did not
meaningfully differ when 80 or 100\% of the data were assigned to OTUs
using the \emph{de novo} methods. Second, the AL method had higher MCC
values than the other methods followed by DGC, AGC, CL, open-reference,
and closed-reference, and SL (Figure 1D). Third, with the possible
exception of the CL method, the MCC values for each of the methods only
demonstrated a small amount of variation between runs of the method with
a different ordering of the input sequences. This indicates that
although there may be variation between executions of the same method,
they produced OTU assignments that were of equal quality. Revisiting the
concept of stability, we question the value of obtaining stable OTUs
when the full dataset is not optimally assigned to OTUs. Our analysis
indicates that the most optimal method for assigning the Canadian soils
sequences to OTUs using a 97\% threshold was the AL method.

>>>>

\textbf{\emph{How do the various methods vary in the quality of their
OTU assignments?}} More important than the stability of OTUs is whether
sequences are assigned to the correct OTUs. A method can generate highly
stable OTUs, but the OTU assignments may be meaningless if they poorly
represent the specified cutoff and the actual distance between the
sequences. To assess the quality of OTU assignments by the various
methods, we made use of the pairwise distance between the unique
sequences to count the number of true positives and negatives and the
number of false positives and negatives for each method and sampling
depth. Counting the frequency of these different classes allowed us to
judge how each method balanced the ratio of true positives and negatives
to false positives and negatives using the MCC. We used the average MCC
value as a measure of a method's quality and its variation as a measure
of its consistency. We made three important observations. First, each of
the \emph{de novo} methods varied in how sensitive their MCC values were
to additional sequences (Figure 1C). The SL and CL methods were the most
sensitive; however, the quality of the OTU assignments did not
meaningfully differ when 80 or 100\% of the data were assigned to OTUs
using the \emph{de novo} methods. Second, the AL method had higher MCC
values than the other methods followed by DGC, AGC, CL, open-reference,
and closed-reference, and SL (Figure 1D). Third, with the possible
exception of the CL method, the MCC values for each of the methods only
demonstrated a small amount of variation between runs of the method with
a different ordering of the input sequences. This indicates that
although there may be variation between executions of the same method,
they produced OTU assignments that were of equal quality. Revisiting the
concept of stability, we question the value of obtaining stable OTUs
when the full dataset is not optimally assigned to OTUs. Our analysis
indicates that the most optimal method for assigning the Canadian soils
sequences to OTUs using a 97\% threshold was the AL method.
